---
title: "Assignment 6 SQL"
author: "Roberta Southwood"
format: html
editor: visual
---

## Part 1: Answer the following questions by running SQL using RStudio

```{r}
# load required packages 

library(DBI) 
library(duckdb)
library(dplyr) 
library(nycflights13)
```

#### **a.** List the first **10 rows** of the `flights` table

```{r}
# connect to duckdb
dbcon <- dbConnect(duckdb::duckdb()) 
dbcon

# check connection 
dbIsValid(dbcon)

# connect to flights.duckdb
url <- "https://data-science-master.github.io/lectures/data/flights.duckdb"

download.file(url, destfile = "./data/flights.duckdb", mode = "wb")

conn <- dbConnect(duckdb(), "./data/flights.duckdb", read_only = TRUE)

# check connection 
dbIsValid(conn)

# list first 10 rows of flights table 
dbReadTable(conn, 'flights') %>% 
  head(10)

```

#### **b.** Find the **number of flights**

i.  in the dataset

ii. for each carrier

```{r}

# SQL command to find number of flights in the dataset
count_flights <- "select count(flight) from flights"

# run query 
dbGetQuery(conn, count_flights)

```

```{r}
# SQL command
count_flights_by_carrier <- "select carrier, count(carrier) 
                from flights
                group by carrier"
# run query 
dbGetQuery(conn, count_flights_by_carrier)

```

#### **c.** Find the **average departure delay** (`dep_delay`)

```{r}
# SQL command
average_dept_delay <- "select avg(dep_delay) from flights"

# run query
dbGetQuery(conn, average_dept_delay)

```

#### **d.** Find the **top 5 destinations** (`dest`) With the most flights.

-   Your output should include:

    -   destination

    -   number of flights Sort from the **largest to smallest**.

```{r}
# SQL command
dest_most_flights <- "select dest, count(dest) from flights
                      group by dest 
                      order by count(dest) desc
                      limit 5"

# run query
dbGetQuery(conn, dest_most_flights)

```

#### **e.** Create a table showing:

-   `carrier`

-   average arrival delay (`arr_delay`)

Sort the results from **largest to smallest delay**

```{r}
# SQL command
arrival_delay <- "select carrier, avg(arr_delay) from flights
                  group by carrier 
                  order by avg(arr_delay) desc"

# run query
dbGetQuery(conn, arrival_delay)

```

> You must close the connection when you are done. Use `dbDisconnect()`

```{r}

# disconnect connection 
dbDisconnect(conn)
dbDisconnect(dbcon)

# check disconnection 
dbIsValid(conn)
dbIsValid(dbcon)

```

## Part 2: Answer the following questions by running SQL using DBeaver with DuckDB  

#### **a.** List the first **10 rows** of the `flights` table

#### **b.** Count the Number of Flights for Each Carrier

Find the **number of flights** for each carrier.

-   Your output should include:

    -   carrier

    -   number of flights Sort from the **largest to smallest**.

#### **c.** Find the **average departure delayby month**

-   Your output should include:

    -   month

    -   average departure delay

#### **d.** Find the **top 10 routes** (`dest`) by:

-   `origin`

-   `dest`

ordered by **number of flights**.

#### **e.** Export the result of part b query to a **CSV file** using the **Export Data** feature in DBeaver.

Submit the exported CSV file along with your homework.

## **Remember to Close the connection When You Are Done**
