---
title: "Assignment 6 SQL"
author: "Roberta Southwood"
format: html
editor: visual
---

## Part 1: Answer the following questions by running SQL using RStudio

```{r}
# load required packages 

library(DBI) 
library(duckdb)
library(dplyr) 
library(nycflights13)
```

#### **a.** List the first **10 rows** of the `flights` table

```{r}
# connect to duckdb
dbcon <- dbConnect(duckdb::duckdb()) 
dbcon

# check connection 
dbIsValid(dbcon)

# connect to flights.duckdb
url <- "https://data-science-master.github.io/lectures/data/flights.duckdb"

download.file(url, destfile = "./data/flights.duckdb", mode = "wb")

conn <- dbConnect(duckdb(), "./data/flights.duckdb", read_only = TRUE)

# check connection 
dbIsValid(conn)

# list first 10 rows of flights table 
dbReadTable(conn, 'flights') %>% 
  head(10)

```

#### **b.** Find the **number of flights**

i.  in the dataset

ii. for each carrier

```{r}

# SQL command to find number of flights in the dataset
count_flights <- "select count(flight) from flights"

# run query 
dbGetQuery(conn, count_flights)

```

```{r}

count_flights_by_carrier <- "select carrier, count(carrier) 
                from flights
                group by carrier"

dbGetQuery(conn, count_flights_by_carrier)

```

#### **c.** Find the **average departure delay** (`dep_delay`)

#### **d.** Find the **top 5 destinations** (`dest`) With the most flights.

-   Your output should include:

    -   destination

    -   number of flights Sort from the **largest to smallest**.

#### **e.** Create a table showing:

-   `carrier`

-   average arrival delay (`arr_delay`)

Sort the results from **largest to smallest delay**

> You must close the connection when you are done. Use `dbDisconnect()`

## Part 2: Answer the following questions by running SQL using DBeaver with DuckDB  
